---
title: "p8105_hw1_kg3006"
output: github_document
date: "2025-09-15"
author: Kevin Guzman (kg3006)
---

# Homework 1: Building Blocks 
## Problem 1 
### Installing Packages Needed 
This downloads the **early_january_weather** dataset from the **moderndive** library

```{r}
library('moderndive') 
data("early_january_weather")
```

### Describing the Dataset
The **early_january_weather** dataset has `r nrow(early_january_weather)` observations and is composed of `r ncol(early_january_weather)` variables/columns. The columns are as following:  
  `r paste(names(early_january_weather), collapse = ", ")`. 
  
Notable important variables include the **year range** (`r range(early_january_weather$year)`), **mean temperature** (`r mean(early_january_weather$temp)`), and the **mean wind speed** (`r mean(early_january_weather$wind_speed)`).

### Making the Scaterplot 
The scatterplot below shows the relationship between **time (hours)** and **temperature**. We will use tidyverse to make the scatterplot. 

```{r}
library(tidyverse)

ggplot(
  data = early_january_weather,
  mapping = aes(x = time_hour, y = temp, color = humid)) +
  geom_point()


```

### Describing The Scatterplot
This scatterplot shows time acrross a couple of weeks in Janurary. Temperatures ranged from `r min(early_january_weather$temp)` to `r max(early_january_weather$temp)` degrees Fahrenheit. The color gradient represents humidity levels, with lighter colors indicating higher humidity. There is an increase in temperature during as Janurary progresses. It also appears that higher temperatures also have higher himidity levels. 
  
### Saving the Scatterplot
```{r}
ggsave("p8105_hw1_kg3006_scatterplot.png")
```

## Problem 2 

In this problem we will create a dataframe ourselves and analyze it. 

### Creating the Dataframe
```{r}
hw1 <- tibble(
  size = rnorm(10, sd = 1), # 10 observatoins created 
  size_greater_than_0 = size > 0, # True if size is greater than 0 
  book_character = c("Alice", "Max", "Kevin", "Jose", "Hannah", "Mirna", "Sam", "Mia", "Omar", "Edwin"),
  occupation = as.factor(c("Doctor", "Professor", "Nurse", "Doctor", "Professor",   "Nurse", "Doctor", "Professor", "Nurse", "Doctor"))) #factor with 3 variables 
```

### Taking the mean of each variable 

#### Size
We will take the mean of the varibles using the pull function. 
```{r}
 hw1 %>% 
  pull(size) %>% 
  mean()
```

The mean for the **size** variable is `r hw1 %>% pull(size) %>% mean()`. This function works. 

#### Logical variable 

Taking the mean of the logical variable **size_greater_than_0**. 

```{r}
 hw1 %>% 
  pull(size_greater_than_0) %>% 
  mean()
```

This also works. The mean here is `r hw1 %>% pull(size_greater_than_0) %>% mean()`. However, since this is treating False as 0 and True as 1 and there are only 10 samples, then the 0.1 means that 1 person is coded as 1 and the rest are coded as 0. 

#### Book characters 

Taking the mean of the **book_character** variable.

```{r}
 hw1 %>% 
  pull(book_character) %>% 
  mean()
```

Taking the mean of a character variable does not work. 

#### Occupation 
Taking the mean of the **occupation** variable.

```{r}
 hw1 %>% 
  pull(occupation) %>% 
  mean()
```

This also does not work for factor variables. 

### Changing the 3 variables into numbers

####Integers to numbers and means
First part of this code changes the integer variables into numbers and the second part tries to analyze there means. 

```{r}
 hw1 <- hw1 %>% 
  mutate(size_greater_than_0 = as.numeric(size_greater_than_0))
  
 hw1 %>% 
  pull(size_greater_than_0) %>% 
  mean()

```

This give the same answer as before because the variables were already read as 1/0. 

#### Characters into numbers
```{r}
 hw1 <- hw1 %>% 
  mutate(book_character = as.numeric(as.factor(book_character)))
  
 hw1 %>% 
  pull(book_character) %>% 
  mean()
```

This also works, and works by converting each character into a number 1 - 10. By adding them all and dividing by 10. We get a mean of `r hw1 %>% pull(book_character) %>% mean()`.

#### Factors into numbers
```{r}
 hw1 <- hw1 %>% 
  mutate(occupation = as.numeric(occupation))
  
 hw1 %>% 
  pull(occupation) %>% 
  mean()
```

This also works, and works by converting each factor into a number 1 - 3. By adding them all and dividing by 10. We get a mean of `r hw1 %>% pull(occupation) %>% mean()`.