---
title: "p8105_hw1_kg3006"
output: github_document
date: "2025-09-15"
author: Kevin Guzman (kg3006)
---

# Homework 1: Building Blocks 
## Problem 1 
### Installing Packages Needed 
This downloads the **early_january_weather** dataset from the **moderndive** library

```{r}
library('moderndive') 
data("early_january_weather")
```

### Describing the Dataset
The **early_january_weather** dataset has `r nrow(early_january_weather)` observations and is composed of `r ncol(early_january_weather)` variables/columns. The columns are as following:  
  `r paste(names(early_january_weather), collapse = ", ")`. 
  
Notable important variables include the **year range** (`r range(early_january_weather$year)`), **mean temperature** (`r mean(early_january_weather$temp)`), and the **mean wind speed** (`r mean(early_january_weather$wind_speed)`).

### Making the Scaterplot 
The scatterplot below shows the relationship between **time (hours)** and **temperature**. We will use tidyverse to make the scatterplot. 

```{r}
library(tidyverse)

ggplot(
  data = early_january_weather,
  mapping = aes(x = time_hour, y = temp, color = humid)) +
  geom_point()


```

### Describing The Scatterplot
This scatterplot shows time acrross a couple of weeks in Janurary. Temperatures ranged from `r min(early_january_weather$temp)` to `r max(early_january_weather$temp)` degrees Fahrenheit. The color gradient represents humidity levels, with lighter colors indicating higher humidity. There is an increase in temperature during as Janurary progresses. It also appears that higher temperatures also have higher himidity levels. 
  
## Problem 2 